---
title: "lab-5"
author: "Alex Steiner"
date: "2024-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Working with matrices, lists, and data frames.

### Question 1.
Assign to the variable ```n_dims``` a single random integer between 3 and 10.
```{r}
n_dims <- round(runif(1,min = 3,max = 10),0)
n_dims
```

Create a vector of consecutive integers from 1 to n_dims^2.

```{r}
v <- c(1:n_dims^2)
v
```

Use the sample function to randomly reshuffle these values.

```{r}
w <- sample(v)
w
```

create a square matrix with these elements.
```{r}
m <- matrix(data = w, nrow = 8, ncol = 8)
```

print out the matrix.
```{r}
print(m)
```

find a function in r to transpose the matrix.

```{r}
uu <- t(m)
```

print it out again and note how it has changed.
```{r}
print(uu)
```
It flips across the diagonal.

calculate the sum and the mean of the elements in the first row and then the last row.

```{r}
sum(uu[1, ])
mean(uu[1, ])
sum(uu[8,])
mean(uu[8,])
```

read about the eigen() function and use it on your matrix

```{r}
euu <- eigen(uu)

#eigen() calculates the eigen values and vectors from a matrix
```

look carefully at the elements of $values and $vectors in the output. What kind of numbers are these?
```{r}
#imaginary?? really weird? 
euu
```


dig in with the typeof() function to figure out their type.

```{r}
typeof(euu)

typeof(euu[[1]]) ## it says complex, but it is a vector

typeof(euu[[2]])  ## it says complex, but it is a matrix
```


if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.



### Question 2. 

Create a list with the following named elements:

my_matrix, which is a 4 x 4 matrix filled with random uniform values
my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_letters, which is a 26-element vector of all the lower-case letters in random order.

```{r}
my_matrix <- matrix(runif(16), nrow = 4)
my <- c(1:100)
my_logical <- my > 50
my_letters <- sample(c(letters))
l <- list(my_matrix, my_logical, my_letters)
```

Then, complete the following steps:
a. create a new list, which has the element[2,2] from the matrix, the second element of my_logical, and the second element of my_letters.

```{r}
nl <- list(l[[1]][2,2], l[[2]][2], l[[3]][2])
```
use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(nl[[1]])
typeof(nl[[2]])
typeof(nl[[3]])
```

combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
av <- c(nl[[1]],nl[[2]],nl[[3]])
alsoav <- unlist(nl)
```


what is the data type of this vector?

```{r}
typeof(av)
str(av)
```
It is a character.

### Question 3.

Create a data frame with the two variables (= columns) and 26 cases (= rows) below:
```{r}
df <- data.frame(matrix(nrow = 26, ncol = 2))
```

a. call the first variable my_unis and fill it with 26 random uniform values from 0 to 10
```{r}
names(df)[1] <- "my_unis"
df$my_unis <- runif(26, min = 0, max = 10)
```
b. call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
names(df)[2] <- "my_letters"
df$my_letters <- sample(LETTERS)
```

c. for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.

```{r}
df[sample(nrow(df), 4), ] <- "NA"
df
```

d.for my_unis, write a single line of R code to identify which rows have the missing values.
```{r}
which(df$my_unis == "NA")

```
e. re-order the entire data frame to arrange the second variable in alphabetical order
```{r}

dfa <- df[order(df$my_letters),]

```

f. calculate the column mean for the first variable.
```{r}
df$my_unis <- as.numeric(df$my_unis)

mean(df$my_unis, na.rm = TRUE)


```









