---
title: "Lab-4"
author: "Alex Steiner"
date: "2024-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Homework lab number 4 
BIO
These problems can be found [here](https://gotellilab.github.io/Bio381/Homeworks/Homework_04.html).

### Question 1
Suppose $x = 1.1$, $a = 2.2$, and $b = 3.3$. Assign each expression to the value of the variable $z$ and print the value stored in $z$.

a. $x^{a^b}$

b. $(x^a)^b$

c. $3x^3+2x^2+1$

```{r}
x <- 1.1
a <- 2.2
b <- 3.3

## a.
z <- x^a^b
print(z)
## b. 
z <- (x^a)^b
print(z)
## c. 
z <- (3*(x^3))+(2*(x^2))+1
print(z)
```

 
### Question 2
Using the ```rep``` and ```seq``` functions, create the following vectors:
 
a. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
b. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
c. $(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)$
```{r}
#a.
l <- seq(from = 1, to = 8, by = 1)
m <- seq(from = 7, to = 1, by = -1)
vector_a <- c(l,m)
print(vector_a)

#b.
n <- seq(from = 1, to = 5, by = 1)
vector_b <- rep(x = n, times = n)
print(vector_b)
#c.
o <- seq(from=5, to=1, by = -1)
vector_c <- rep(x = o, time = n)
print(vector_c)
#is there convention to not assign values to "o", as it looks like a 0?
```
  
# Question 3
Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (```asin()```, ```acos()```, or ```atan()```), convert these numbers into polar coordinates.
```{r}

set.seed(1137) #1/137 --look it up
cart_Coords <- runif(2) #make random coords
print(cart_Coords)

r = sqrt((cart_Coords[1]^2) + cart_Coords[2]^2) ##calc radius
theta <- atan(cart_Coords[2]/cart_Coords[1]) ##calc theta(angle)

polar_Coords <- c(r, theta)
print(polar_Coords)
```
  
### Question 4.
Create a vector ```queue <- c("sheep", "fox", "owl", "ant")``` where ```queue``` represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update ```queue``` as:
```{r}

queue <- c("sheep", "fox", "owl", "ant")

#a. the serpent arrives and gets in line;
queue_a <- c(queue, "serpent")

#b. the sheep enters the ark;
queue_b <- queue_a[2:5]

#c. the donkey arrives and talks his way to the front of the line;
queue_c <- c("donkey", queue_b)

#d. the serpent gets impatient and leaves;
queue_d <- queue_c[-5]

#e. the owl gets bored and leaves;
queue_e <- queue_d[-4]

#f. the aphid arrives and the ant invites him to cut in line.
queue_f <- c(queue_e, "ant", "aphid")

#g. Finally, determine the position of the aphid in the line.
queue_g <- which(queue_f == "aphid")

```
### Question 5. 
Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7. 
```{r} 
hundo <- seq(from=1, to = 100, by = 1) ##make vector of 1 to 100
## select the indeces in hundo which have a remainder when divided by 2 
divide2 <- hundo[which((hundo%%2)>0)] 
divide3 <- divide2[which((divide2%%3)>0)]
divide7 <- divide3[which((divide3%%7)>0)]

print(divide7)

##Hi George, which of these methods is better practice? 
notdivis <- hundo[which((hundo%%2)>0 & (hundo%%3)>0 & (hundo%%7)>0)]
print(notdivis)


```

